### 2020-05-14 ctapc/paper.tex
未來工作

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{The Design of Shallow State Blockchain*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
Cryptocurrency based on blockchain is a peer to peer trading system.
In this decentralized system, each node maintain a complete ledger which records entire transaction history of all account.
The system has the characteristics that it is very hard to tamper with the data.
However, such characteristic only makes the data increase but do not decrease.
Years after years, requirement to single node will get higher and higher.
Once some entities get overwhelmed, the degree of centralization of this blockchain system will gradually rise

Dealing with the issue, stateless blockchain is a promising approach.
To be more specific, in this system, each transaction is attached with proof of validity.
With this approach, there is no need for a node to access the disk to search past information of the ledger.
Instead, node only verifies the proof attached to each transaction.
Therefore, it will significantly reduce the space of data storage for each node.

In this paper, we will discuss the following two points.
First, it only need to cache part of information instead of attaching proof to each transaction.
It is highly possible to infer the validity of a transaction from other valid transaction appearing recently.
Second, when blockchain forks, it have to go back to the bifurcation point and create new cache of new blocks.
To reach the purpose, we propose an immutable data structure so that cache can be rapidly computed under any situation.
Finally, we conduct the experiments with different workload to understand the performance of different cache strategies and data structures
\end{abstract}

\begin{IEEEkeywords}
blockchain, stateless blockchain, data structure
\end{IEEEkeywords}

\section{Introduction}
In 2008, Satoshi Nakamoto proposed the white paper of bitcoin. It is a new type of cryptocurrency system based on blockchain. In this system, each node maintain a complete ledger in order to attain the high degree of decentralization of entire system. However, as the data size of ledger got bigger and bigger, storage space of node increased yearly. Ten years later, it costs hundreds of GB storage space for a node to run some well-known cryptocurrency system, such as Bitcoin and Ethereum. Thus, it gradually make the nodes maintained by person overload. If we settle for second best by saving the information needed for verifying new block, such as UTXO in Bitcoin and world state in Ethereum, it still needs several or tens of GB for disk storage place. With this method, not only the storage space of disk is requested but also  I/O speed. Furthermore, it is too slow for traditional disk to access randomly. The speed of verifying a block on Ethereum can't keep up with the generation speed of blocks. It force the node maintainer on Ethereum to buy expensive solid-state disk, and it further cause the increase in operation cost of node.

To deal with the problem above, cryptocurrency communities start trying several solutions, and stateless blockchain is prospective one of them. It only need to store block header for each node on stateless blockchain. Nevertheless, each transaction must be attached with proof of validity. This approach significantly reduces the requirement of storage space. Furthermore, it is no need to search the validity of transaction on disk; namely, it drastically reduces the I/O time.

This paper proposes an imrpovement to stateless blockchain. We let the node on stateless blockchain reach a consensus to maintain a cache. If the payer of transaction appears in recent transaction, it is highly possible to infer the validity from cache. Hence, due to the omission of proof of validity, it reduces the size of transaction and the Internet traffic while broadcasting. However, how to maintain cache effectively becomes a new problem. When a new block append to the blockchain, there will be a correspondent cache. Nevertheless, there will also exist overlap on the blocks nearby. Therefore, we design persisted data structure for different cache strategies, such as the nearest k blocks and LRU, and we conduct an experiment to test their performance.

The remaining sections are arranged as follows. Section II narratives some related works, whereas section III describes the background of this paper and preliminary knowledge; section IV illuminates the design of stateless blockchain, including cache strategies and data structures used by cache; section V shows the design and results of experiment; section VI discusses future extensions to stateless blockchain, and concluding remarks are found in section VII.

\section{design}
In order to reduce occupied space of the disk, it is necessary for a stateless blockchain to attach a proof to each transaction. However, this move will increase Internet traffic. Furthermore, to reduce the time of accessing disk, it will cost computational power for CPU to verify the proof.
Compared to normal blockchain, stateless blockchain made some trade-offs. Regarding the purpose of shallow-state blockchain, the trade-offs are no longer all or nothing, which means that all transactions are attached with proofs or none of transactions are attached with proofs. It makes the degree of state storage adjustable.

\subsection{shallow-state blockchain}
In stateless blockchain, if payers and payees in some transactions are the same, the proofs in these transactions will be identical. It is obvious that the same proofs can be omitted in a block. Extended by this idea, if we cache the accounts’ information appearing in recent transactions, there is no need to verify the proof again when the accounts make transactions. To go a step further, if nodes on the Internet follow the same set of rules to record cache, it will reach a consensus that node knows whether to attach the proof. While broadcasting the block, nodes can get rid of unnecessary proofs, and then reduce internet traffic.

\subsection{cache design}
The problems mentioned above can be avoided by a simple design principle. Namely, each block has its own cache, and the content of cache depends on which chain the block is situated. Therefore, we turn a tree structure into a list, and it is definitely identical for the block on different nodes. If every node uses the same deterministic algorithm to compute the cache from this list, it can ensure that every node gets identical results of cache after verifying the same block.

\subsection{persistent cache}
In this part, we take another approach: every block retains its own cache. When a new block is about to append to the chain, it can directly use the cache of the previous block. Thus, there is no need to compute cache again. In other words, we adopt fully persistent data structure to store cache. Everytime the chain gets a new block, it creates a new version of cache.
In this method, we have to delete the cache of older blocks, such as the block 20 blocks far from the newest one, in order to limit the size of cache to a certain range. Otherwise, it will run out of memory, and it will violate our design principle.
Adopting this method leads to a new problem. If we duplicate the cache of the previous block, the space of cache will be proportional to the cache numbers that we didn’t delete. However, there is a high similarity of the cache in adjacent blocks. If we use appropriate data structure to have the cache shared by adjacent blocks, it will effectively increase space usage.

\subsection{LRU}
LRU is an acronym for “least recently used”. In this strategy, the size of cache is fixed. If a cache is full, it will discard a piece of data before inserting new data. The least recently used data are chosen by LRU strategy, and it will be discarded.
The reasons why LRU has higher hit rate than FIFO are revealed by the history data of Ethereum. First, after an account spends money, it is highly possible that it will spend money again. Second, after an account receives money,  it is highly possible that it will spend money right away. Because LRU updates the using time of data, it is able to keep caching the data appearing frequently. However, FIFO only looks at the coming time of data. Even if some data appear frequently, they will be discarded sooner or later.
From an abstract view, LRU is a data structure that supports two kinds of interface.
\begin{itemize}
\item get(key)
\item put(key, value)
\end{itemize}
When executing get(key), LRU will determine whether the key exists. If the key exists, it will return the value and update its using time.
When executing put(key, value), LRU will determine whether the space of LRU is full. If the space of LRU isn’t full, the key-value pair will be inserted to the space directly.  If the space of LRU is full, LRU will find out the least recently used one in the cache and abandon it, and then insert the new key-value pair to it. The using time of the new key-value pair should be the latest.
Corresponding to the situation of stateless blockchain, everytime when a new block appends to the chain, we have to compute the cache of the new block. In this time, a series of read and modify execution of accounts will turn into get and put. Key will be the address of account, and value will be the state of account. Finally, the underlying data structure of LRU will take the correspondent action.

\subsection{persistent LRU algorithm}
Before discussing persistent LRU algorithms, we first observe how to efficiently implement LRU cache on software. Judging from several popular LRU libraries on github, interior data structures are all combinations of hash table and doubly linked list.
While executing get, it will get the data by the pointer pointing to the node of list. Then, the node will move to the head of the list. While executing put, if cache hits, the value of node will be updated, and the node will move to the head of the list. If cache is not full, the cache value will be added to the head and the pointer will be put in the hash table. Instead, if the cache is full, pull out the tail node of the doubly linked list, and delete the old key in the hash table. Then, add cache value to the head and put the pointer in the hash table.
We observed that there are two kinds of information that need to be recorded by LRU.
\begin{enumerate}
  \item find the value from key (key-value pair)
  \item the order information of all keys
\end{enumerate}
In the implementation of combinations of hash table and doubly linked list, hash table is responsible for (1), and doubly linked list is responsible for (2). We observed that a doubly linked list records the order relationship naturally, and it doesn't even need to record the definite using time.
There is one simple idea that we directly combine persistent-versioned alternatives of a hash table and a doubly linked list, and then we get a persistent LRU. Although there are mature alternatives for hash tables, it is hard to find such a thing for doubly linked lists.
Thus, we need to use other persistent and efficient data structures to replace doubly linked lists. The things that an abstract doubly linked list can do are as follows.
\begin{itemize}
\item Update the using time of node to latest
\item Delete the least recently used node
\item Insert new node and update the using time of the node to latest
\end{itemize}
Following we will discuss how to use red-black tree to complete the above tasks and introduce our design of order tree, which is more efficient than red-black tree.

\subsection{hash table + red-black tree}
We tried to use red-black to record the order information.
First, set a unique timestamp for every account.
This timestamp is easy to access. For instance,
it could be $block\_height * max\_tx\_in\_one\_block+ tx\_number$.
Afterwards, build a red-black tree whose key is timestamp and value is account’s information. About hash table, the account's address is set as key and the correspondent timestamp is set as value.
While executing get, it first gets the timestamp from the account's address in the hash table, and then gets the account’s information from the timestamp in the red-black tree. For example, in Fig 3.6, we first get the account’s timestamp, 243, from the hash table, and search 243 in red-black tree.
After searching, it needs to update the using time. The specific operations are that modify the timestamp corresponding to the address in the hash table and remove the original node in red-black tree. Then, set the new timestamp as key. The put operation is similar to get operation. However, it needs to modify the account's information. 
Fig 3.6 shows the sharing structure of the persistent red-black tree. In this figure, the cache size is 8. In state 1, there are only 7 data. After that, state 2 insert 14 in it, and there are 8 data. State 3 insert 15 to state 2, and the cache is full. Therefore, it has to delete the data of minimum timestamp; that is, the red 7 at lower left corner will be deleted.

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}