### 2020-06-14 ntuvars.tex

% author: Tz-Huan Huang [http://www.csie.ntu.edu.tw/~tzhuan]

% ----------------------------------------------------------------------------
% "THE CHOCOLATE-WARE LICENSE":
% Tz-Huan Huang wrote this file. As long as you retain this notice you
% can do whatever you want with this stuff. If we meet some day, and you think
% this stuff is worth it, you can buy me a chocolate in return Tz-Huan Huang
% ----------------------------------------------------------------------------

% Syntax: \var{English}{Chinese}
\university{National Taiwan University}{國立臺灣大學}
\college{College of Electrical Engineering and Computer Science}{電機資訊學院}
\institute{Department of Computer Science and Information Engineering}{資訊工程學系}
\title{無狀態區塊鏈及運用 LRU 快取策略的評估}{Structural Assessment on Stateless Blockchain Improvements with
LRU-oriented Cache Replacement Policy}
\author{TSUN-CHE TSAI}{蔡存哲}
\studentid{R05922086}
\advisor{Shih-Wei Liao, Ph.D.}{廖世偉 博士}
\defenseyear{2020}{109}
\defensemonth{May}{5}
\defenseday{1}
\doi{doi:10.6342/NTU2017XXXXX}
\keywords{keyword}{關鍵字}